Stream Mining Algorithms for Sensor Data Classification

### 2015-05-31 latex/paper-draft.tex
Revision of grammar, sentence structure, typos, etc. throughout first 60% of paper.

%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\usepackage[caption=false,font=footnotesize]{subfig}
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

@@

		\begin{abstract}
			%\boldmath
			Learning from data streams is a research area of increasing importance \cite{evaluation_stream_mining}. Nowadays, several stream learning algorithms have been developed. Most of them learn decision tree models that continuously evolve over time, such as Very Fast Decision Trees (VFDT). However, it is not natural to use decision trees when learning has to be continuous --- which it frequently does due to concept drift --- because it is costly to adjust the tree's shape with new data. We thus propose two efficient incremental learning algorithms based on Neural networks, Sliding Mini-batch Neural Networks and Sliding Ensemble Neural Networks, which avoid this problem. The results of our experiment on the Dodgers Loop Sensor Dataset indicate that the performances of our algorithms converge to that of popular batch learners.

@@

		Learning from data streams is a research area of increasing importance \cite{evaluation_stream_mining}. Many datasets are no longer static, instead they continuously get updated, sometimes with millions or even billions of new records per day. In most cases, this data arrives in a stream or in streams, and if it is not processed immediately or stored, then it will be lost forever. Moreover, the data arrives so rapidly that it is usually not feasible to store it all. To extract useful knowledge from this tremendously valuable data, data stream mining became popular.

		Data stream mining is the process of extracting knowledge from an ordered sequence of data instances. One goal of data stream mining is to predict the class of new data instances based on previous instances in the data stream (classification). In many applications, data analysis can only be performed in one pass over the data, storing at most a small fraction of data points. Because of this, traditional batch learning algorithms are often unsuitable for data stream mining.  We thus need algorithms that can summarize information about prior data instances without keeping all of them in memory.

		Several attempts have been made to modify the existing batch learners for data stream learning. According to Gama et al. \cite{evaluation_stream_mining}, most of them use decision trees ``that continuously evolve over time, taking into account that the environment is non-stationary and computational resources are limited.'' Examples of these algorithms are Very Fast Decision Trees (VFDT) \cite{VFDT} and Concept adapting Very Fast Decision Trees (CVFDT) \cite{CVFDT}.

@@

		We build two models based on mini-batches of Neural Networks and sliding windows.
		The first model \emph{\textbf{S}liding \textbf{M}ini-batch  \textbf{N}eural \textbf{N}etworks} (SMNN) works by passing the final weights from the trained neural network on one sliding window to be the initial weights at the subsequent window. The second model \emph{\textbf{S}liding \textbf{E}nsemble  \textbf{N}eural \textbf{N}etwork} (SENN) works by training one neural network -- completely independently -- in each sliding window and taking a majority vote from all neural networks in the ensemble.

@@

		The rest of this paper is organized as follows: in Section~\ref{sec:related-work}, we discuss the related work. In Section~\ref{sec:models}, we propose our methods for mining streaming data. In Section \ref{sec:experimental-design}, we discuss the data pre-processing and the experimental design. In Section~\ref{sec:experiments-and-results}, we present the results and compare them with the popular stream mining algorithms. In Section~\ref{sec:conclusion}, we conclude with remarks, discussion of the limitation of the experiment and suggestions for further work.

@@

		Stream data mining have become prevalent only in the last few decades; however, many literature already proposed different algorithms for stream data mining. Based on paper ``Introduction to stream: An Extensible Framework for Data Stream Clustering Research with R'' \cite{Rstream}, several classification methods suitable for data streams have been developed. Examples are Very Fast Decision Trees (VFDT) \cite{VFDT} based on Hoeffding trees, Online Information Network (OLIN)  \cite{last}, On-demand Classification \cite{Aggarwal04} based on micro-clusters found with the data-stream, and clustering algorithm CluStream \cite{ggarwal03}. There are some other methods we found online such as UFFT (ultra fast forest trees) \cite{gama} and CBDT (concept based decision tree) \cite{CBDT}.

@@

		The traditional decision tree classifier uses a divide-and-conquer approach. Typically, it needs to scan the whole dataset once to grow one level (excluding elements that have already reached a leaf at a lower level). After the scan, it finds the best attribute for splitting the data at each node by information gain or etc. To develop one decision tree, multiple scans of the whole dataset are required as the following figure indicates:
		\begin{figure}[H]
			\centering
			\includegraphics[width=\linewidth]{figures/decision_tree}
			\caption{An example of decision tree}
			\label{fig:tree}
		\end{figure}

		As we can see, decision trees need to scan the whole dataset to grow one level (if no leaves have been formed yet). However, we cannot keep the entire dataset in memory for typical stream mining datasets. Therefore, traditional decision trees algorithms don't suit the job of stream mining. An algorithm that can learn \emph{incrementally} is required.

		The VFDT overcomes this limitation by building each node of the decision tree based on a small amount of data instances. According to Domingos \cite{VFDT}, the VFDT works as the following: ``given a stream of examples, the first ones will be used to choose the root test; once the root attribute is chosen, the succeeding examples will be passed down to the corresponding leaves and used to choose the appropriate attributes there, and so on recursively." An illustration of how VFDT is built is shown as Figure~\ref{fig:VFDT}.

		\begin{figure}[H]
			\centering
			\includegraphics[width=\linewidth]{figures/vfdt1} 
			\caption{An illustration of VFDT}
			\label{fig:VFDT}
		\end{figure}
		To decide how many examples (the size of $n$ in the above Figure) are necessary at each node, the Hoeffding bound is used. With the hoeffding bound, the VFDT can guarantee that the performance of Hoeffding tree converges to that of a batch learner decision tree \cite{VFDT}.

@@

		Although the performance of the VFDT converges to that of a batch learning tree, this data structure is not ``stable'' and only works when there is no concept drift. Subtle changes of the distribution of the data can cause the Hoeffding trees to fail. For example, in Figure~\ref{fig:VFDT}, suppose that the first $n$ examples choose Q1 as the root attribute, and we kept growing trees until we reach level 1000. Suddenly, the distribution of the data changed, choosing Q500 as the root attribute will give a much better prediction as choosing Q1. Then the tree has to rebuild again, but we can't retrieve the data passed in the stream, and we might not have enough data to rebuild another tree.

		To mitigate the issue of concept-drift, the same authors introduced Concept-adapting Very Fast Decision Trees (CVFDT), which adapt better to concept drift by monitoring changes of information gain for attributes and generating alternate subtrees when needed. However, it still can't solve the issue completely because the algorithm only monitors the leaves.

@@

		Many researchers present algorithms based on incremental neural networks, such as the paper \cite{NN1,NN2,NN3,NN4} which use incremental neural networks to training data one by one, and update the weight every time after scanning a data points. The shortcoming of this incremental neural networks is that the result is highly depends on the randomized initial weight since we only pass the data once. As we know, in batch learning, we pass the data in many epochs to avoid the effect of randomized initial weight. This is impossible in incremental neural networks since the data only got passed once, and we can't retrieve the lost data.

		Other researchers present methods based on ensemble classifiers, such as paper \cite{ensemble1} propose a general framework for mining concept-drifting data streams using weighted ensemble classifiers. The authors train an ensemble of classification models from sequential chunks of the data stream. However, building a classifier independently on each chunk can be inefficient since the data from each chunk are usually dependent.

		Overall, most of stream mining algorithms involve some stream summarization. There are different ways to summarize data streams. Some methods will only keep a portion of the data in streams by using sampling. Such approaches including Reservoir Sampling \cite{vitter}, Concise Sampling \cite{Gibbons}, Chain Sampling \cite{Babcock}, Order Statistics with sampling approximate quantiles \cite{Manku}, and Sampling from a Moving Window, such as sliding windows. Another common method for processing the stream data is filtering, such as bloom Filters. The third common method is to construct a probabilistic summary of the data with hash functions and randomized algorithms to get approximation.

		The models present in this paper overcome the disadvantages of incremental neural networks and ensemble classifiers.  Our method of SMNN learn the stream data in mini-batches with sliding windows. In each mini-batch, we run the algorithm in many epochs to reduce the effect of randomized initial weight. Moreover, we pass the weight from one mini-batch to the other mini-batch in order to obtain a better weight. The method of SENN trains multiple neural networks on sliding windows and uses plurality voting on the collection of neural networks to classify new data. Moreover, whenever this is no concept drift, we pass the weights learned from a sliding window to the subsequent sliding window and use them as the initial weights of the next neural network. By doing so, the training  is much faster as compared to using random weights every time. We will also show that the classifiers are diverse enough to give a good prediction.

		\section{Sliding Neural Network Algorithms} 
		\label{sec:models}
		As discussed in Section \ref{sec:related-work}, each stream mining algorithm usually involves some summarization method. In both of our models,  we used the \emph{sliding windows} to summarize the data streams.

		Sampling from a Moving Window is to maintain a \emph{sliding window} of the most recently arrived data. In other words, the window has a fixed size $L$ and it works as a first in first out queue. we chose sliding windows to summarize and sample our data, as it is simple and quite appropriate for the task of classification.

@@

		\subsection{Sliding Mini-Batch Neural Networks}
		The first model we came up for mining a stream of data is the Sliding Mini-batch Neural Networks. As the name of Sliding Mini-batch Neural Networks indicated, what we did is to ``pass'' the weights of neural networks. In other words, we passed the final weights of a trained neural network on a sliding window and used them as the initial weights for training a neural network on the next sliding window.

		In our model, we combined the neural networks with sliding windows to train the data. Suppose the training dataset has the size $n$. Let us fix $L$ as the size of sliding windows and $d$ as the length of the step we slide each time, then we will have $(n-L+d)$ \textbf{mod} $d$ ($(n-L+d)\%d$) sliding windows in total. We train our model in the following steps: 
		\begin{enumerate}
			\item We randomly initialize a weight matrix $\textbf{W}_1$
			\item Train the neural network with random assigned weight on the $L$ data points in sliding window for up to $m$ epochs. Obtain $\textbf{W}_1$ which is the final weight matrix in the neural network.
			\footnote{An epoch is a single pass through the entire training set} 
			\item Train the neural network with $\textbf{W}_1$ as starting weights on the $L$ data points in sliding window 2 for $m$ epochs. Obtain $\textbf{W}_2$, the final weights of the neural network.\\
			$\ldots$
			\item Repeat until we reach the last sliding window and obtain the final weights $\textbf{W}_{(n-L+d)\%d}$.
		\end{enumerate}
		A graph of the Sliding Mini-batch Neural Networks is as follows:
		\begin{figure}[H]
			\centering
			\includegraphics[width=1\linewidth]{figures/CNN}
			\caption{Sliding Mini-Batch Neural Networks}
			\label{fig:minibatchANN}
		\end{figure}
		After passing all the data in sliding windows for training, the neural network model is ready for classification. The weight matrix $\textbf{W}_{(n-L+d)\%d}$ will be the weights of connections for the final neural network used to classify testing data.

		As we discussed in section \ref{sec:related-work}, SMNN optimizes the incremental neural networks algorithm since the initialized weight is adjusted constantly by the $L$ data instances in one sliding window. By feeding the data instances in one sliding window in many epochs, the order of feeding the data become less important. Therefore, the performance of SMNN is less random than incremental neural networks algorithm. Therefore, SMNN has smaller bias and variance in classification.

		\subsection{Sliding Ensemble Neural Networks}
		The second model we came up with is based on ensemble of Neural networks with sliding windows. As shown in paper \cite{ensemble1}, ensemble of classifiers usually works better than a single classifier in mining data streams. Therefore, we decided to build one neural network on each sliding window and take a majority vote. However, most of the time, the data instances in data stream are dependent. For example, the temperatures measured by the sensor are usually dependent on or correlated to last temperature measure.

		Because data instances in a data stream are usually dependent, it is nature to assume that the data instances in a sliding window are correlated to those in the previous sliding window.  Due to the dependency of data instances between different sliding windows, we thus concluded that it is inefficient to build a neural network on each sliding window independently.

		We thus further improved the model by passing the weights from one neural network to the succeeding neural network when there is no concept drift. By doing so, the training time would be reduced when using ``good weights'' as the initialization weights.  Therefore, we came up with the second idea called \emph{Sliding Ensemble Neural Networks}. This model works as the following figure shows.

@@

		The building of this model is counter-intuitive because for an ensemble of classifiers to be more accurate than any of the individual members, it is necessary and sufficient that the classifiers are accurate and diverse \cite{Hansen}. An accurate classifier is one that has an error rate of better than 50\%. Two classifiers are diverse if they don't make errors on the same new data points \cite{Dietterich}. We therefore believe that there is a trade-off between the accuracy and diversity in the ensemble. For our model, by passing the weight when there is no concept drift, we eventually trained the classifier on more data points which increases the accuracy of a single classifier in the ensemble. On the other hand, since the classifiers are built from similar data with related weight, we lose the diversity.

		However, the diversity issue can be solved by the models picking. The models picking means we only pick a constant $K$ classifiers to be in the ensemble. With the principle that we only pick classifiers who are diverse enough to each other, we can have the classifiers in the ensemble more accurate and diverse.

		The models picking also will help to reduce the memory need to run the algorithm. One of the main requirements for a stream mining algorithm was \emph{constant memory}. This is not an issue for the Sliding Mini-batch Neural Networks algorithm, since the memory used is constant and only depends on the number of weights of the neural network. The same weights matrix is reused when moving from one sliding window to the next, so memory use does not change.

		However, it is necessary to consider this issue for Sliding Ensemble Neural Networks, since it involves building a collection of classifiers. Of course, one cannot keep adding models (one per sliding windows) indefinitely. Therefore, we fix the maximum number of classifiers that a collection will use as a constant, denoted by $K$. This number is generally much much lower than $n$, the total number of sliding windows.

		Models (neural networks) are added to the collection until its size reaches $K$. At this point, the collection is full and we need to determine a way to \emph{continue learning}, so that the algorithm can adapt to the changing data (in the case of concept drift). In order to do so, some models will have to be dropped from the collection so that new models can be included. The principle of models picking is that the models in the ensemble are diverse in order to give better prediction.

		We introduce three methods that we designed to pick the models.

		\begin{enumerate}
			\item The first method is to use a first in first out (FIFO) queue. The oldest model in the collection is dropped to make room for the new model. This is the simplest method and it allows the learning algorithm to adapt very quickly to changing concepts. However, it is susceptible to develop bias quickly if presented with a lot of unrepresentative training data.

			\item The second method involves random replacement with a non-uniform distribution. There are many non-uniform distributions that would do, the point is to make it such that models at the end of the collection rarely get replaced (long-term memory of the classifier), whereas models at the beginning of the collection frequently get replaced (ability to adapt quickly to change).

			For example, here is a valid way to do this. First, generate a random integer $r$ chosen according to the exponential distribution $\exp(\lambda = 0.1 K)$. Then, let $n = min(floor(r), K)$. We would then discard the model at position $n$ (counting from 0) in our ordered collection of models and replace it with a new model.

			\item The third method is to use error correlation rate to pick up models in order to achieve the diversity in the ensemble. We want to replace those models who give the errors on the same data points. To decide how two models are diverse from each other, we compute the error correlation of all models on a testing dataset. For example, we use all data in one sliding window as the testing data. For each model, we print all errors made by this model. Therefore, we obtain a set of error vectors $X_1, \ldots, X_m \in \{0,1\}^n$, where $m$ is the number of models in the ensemble and $n$ is the number of data instances in the testing set. The attributes of the vectors are either 1 or 0, with 1 indicating a classification error and 0 indicating success. In order to measure the degree of diversity among the models, we then compute the pairwise dot product of all error vectors. In other words, we construct the $m \times m$ matrix $A = (a_{i,j})$ where $a_{i,j} = X_i \cdot X_j$. Small values of $A$ are thus considered ideal, as they indicate strong pairwise independence between models, whereas large values of $A$ are undesirable. Note that $A$ is the Gram matrix of $X_1, \ldots, X_m$. Models who tends to make the mistakes on the same data are undesirable in an ensemble \cite{Dietterich} and they can readily be identified from the matrix $A$.
		\end{enumerate}

@@

		\section{Experimental Design}
		\label{sec:experimental-design}
		\subsection{Dataset}
		The dataset we choose for our experiment is the Dodgers Loop Sensor Data Set from the UCI repository \cite{dataset}. It comes with two file: Dodgers.data and Dodgers.events. Dodgers.data contains 50,400 instances with 3 attributes (date, time, count of cars). Dodgers.events contains 81 instances with 6 attributes representing  Date, Begin event time, End event time, Game attendance, Away team, and W/L score.

@@

		The observations of this dataset were taken over 25 weeks with count aggregates in every 5 minutes. Thus, 288 time data instances were generated per day with each data instance representing count of cars in a 5 minutes slot. The goal of this dataset is to predict the presence of a  game at Dodgers stadium.

@@

		Figure \ref{fig:typical-week} is the plot of a typical week of the Dodgers Loop Dataset. The blue lines indicate the beginning of games and the red lines indicate the end of games.

@@

		Instead of predicting if individual data points fall within the window of a Dodgers game (between a blue line and a red line on the same day), we chose to work on the simpler task of predicting whether or not there was a Dodgers game held on any given day.
		Therefore, we aggregate the data into days containing 288 integers each (as there are 288 five-minute intervals in a day). We then labelled days on which a Dodgers game was held with ``1'' and a day without a game as ``0''.

>>>>

%\usepackage[caption=false]{caption}
\usepackage[font=footnotesize]{subfig}
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\usepackage[caption=false,font=footnotesize]{subfig}
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

@@

		\begin{abstract}
			%\boldmath
			Learning from data streams is a research area of increasing importance \cite{evaluation_stream_mining}. Nowadays, several stream mining algorithms have been developed. Most of them learn decision tree models that continuously evolve over time, such as Very Fast Decision Trees (VFDT). However, it is not natural to use decision trees when learning has to be continuous --- which it frequently does due to concept drift --- because it is costly to adjust the tree's shape with new data. We thus propose two efficient incremental learning algorithms based on neural networks, Sliding Mini-batch Neural Networks and Sliding Ensemble Neural Networks, which avoid this problem. The results of our experiment on the Dodgers Loop Sensor Dataset indicate that the performances of our algorithms converge to that of popular batch learners.

@@

		Learning from data streams is a research area of increasing importance \cite{evaluation_stream_mining}. Many datasets are no longer static, instead they continuously get updated, sometimes with millions or even billions of new records per day. In most cases, this data arrives in streams, and if it is not processed immediately or stored, it will be lost forever. Moreover, the data arrives so rapidly that it is usually not feasible to store it all. To extract useful knowledge from this tremendously valuable data, stream mining became popular.

		Data stream mining is the process of extracting knowledge from an ordered sequence of data instances. One goal of data stream mining is to predict the class of new data instances based on previous instances in the data stream (classification). In many applications, data analysis can only be performed in one pass over the data due to the sheer size of the dataset. Because of this, traditional batch learning algorithms are often unsuitable for data stream mining.  We thus need algorithms that can summarize information about prior data instances without keeping all of them in memory.

		Several attempts have been made to modify the existing batch learners for data stream learning. According to Gama et al. \cite{evaluation_stream_mining}, most of them use decision trees ``that continuously evolve over time, taking into account that the environment is non-stationary and computational resources are limited.'' Examples of these algorithms are Very Fast Decision Trees (VFDT) \cite{VFDT} and Concept-adapting Very Fast Decision Trees (CVFDT) \cite{CVFDT}.

@@

		In this paper, we build two models based on mini-batches of neural networks and sliding windows.
		The first model, \emph{\textbf{S}liding \textbf{M}ini-batch  \textbf{N}eural \textbf{N}etworks} (SMNN), works by passing the final weights from the trained neural network on one sliding window to be the initial weights at the subsequent window. The second model ,z\emph{\textbf{S}liding \textbf{E}nsemble  \textbf{N}eural \textbf{N}etwork} (SENN), works by training one neural network -- completely independently -- in each sliding window and taking a majority vote from all neural networks in the ensemble.

@@

		The rest of this paper is organized as follows: in Section~\ref{sec:related-work}, we discuss the related work. In Section~\ref{sec:models}, we propose our methods for mining streaming data. In Section \ref{sec:experimental-design}, we discuss the data pre-processing and the experimental design. In Section~\ref{sec:experiments-and-results}, we present the results and compare them with the popular stream mining algorithms. In Section~\ref{sec:conclusion}, we conclude with remarks, discussion of the limitations of the experiment and suggestions for further work.

@@

		Stream data mining has become prevalent in the last two decades and many different algorithms for stream data mining have already been proposed in the literature. Based on the paper ``Introduction to stream: An Extensible Framework for Data Stream Clustering Research with R'' \cite{Rstream}, several classification methods suitable for data streams have been developed. Examples are Very Fast Decision Trees (VFDT) \cite{VFDT} based on Hoeffding trees, Online Information Network (OLIN)  \cite{last}, On-demand Classification \cite{Aggarwal04} based on micro-clusters found with the data-stream, and the clustering algorithm CluStream \cite{ggarwal03}. There are many other popular algorithms, such as UFFT (ultra fast forest trees) \cite{gama} and CBDT (concept based decision tree) \cite{CBDT}, for example.

@@

		The traditional decision tree classifier uses a divide-and-conquer approach. Typically, it needs to scan the whole dataset once to grow one level (excluding elements that have already reached a leaf at a lower level). After the scan, it finds the best attribute for splitting the data at each node by information gain or Gini index or other criteria. To develop one decision tree, multiple scans of the whole dataset are required as the following figure indicates:
		\begin{figure}[H]
			\centering
			\includegraphics[width=\linewidth]{figures/decision_tree}
			\caption{An example of decision tree}
			\label{fig:tree}
		\end{figure}

		As we can see, decision trees need to scan the whole dataset to grow one level (if no leaves have been formed yet). However, we cannot keep the entire dataset in memory for typical stream mining datasets. Therefore, traditional decision tree algorithms don't suit the job of stream mining. An algorithm that can learn \emph{incrementally} is required.

		The VFDT overcomes this limitation by building each node of the decision tree based on a small amount of data instances. According to Domingos \cite{VFDT}, the VFDT works as follows: ``given a stream of examples, the first ones will be used to choose the root test; once the root attribute is chosen, the succeeding examples will be passed down to the corresponding leaves and used to choose the appropriate attributes there, and so on recursively." An illustration of how VFDT is built is shown as Figure~\ref{fig:VFDT}.

		\begin{figure}[H]
			\centering
			\includegraphics[width=\linewidth]{figures/vfdt1} 
			\caption{An illustration of VFDT}
			\label{fig:VFDT}
		\end{figure}
		To decide how many examples (the size of $n$ in the above Figure) are necessary at each node, the Hoeffding bound is used. With the Hoeffding bound, the VFDT can guarantee that the performance of Hoeffding trees converges to that of a batch learner decision tree \cite{VFDT}.

@@

		Although the performance of the VFDT converges to that of a batch learning tree, this data structure is not ``stable'' and only works when there is no concept drift. Subtle changes of the data distribution can cause the Hoeffding trees to fail. For example, in Figure~\ref{fig:VFDT}, suppose that Q1 was chosen as the root attribute after the first $n$ examples and trees were grown until reaching level 1000. Next, assuming a sudden change in the data distribution (due to concept drift), it would be possible that choosing a new root attribute, such as Q500 instead of Q1, would give a much better prediction. In this case, the tree would have to be rebuilt completely, but this is generally impossible since we can't retrieve the earlier data passed in the stream.

		To mitigate the issue of concept drift, the same authors introduced Concept-adapting Very Fast Decision Trees (CVFDT), which adapt better to concept drift by monitoring changes of information gain for attributes and generating alternate subtrees when needed. However, it still can't solve the issue completely because the algorithm only monitors the leaves.

@@

		Many researchers have proposed algorithms based on incremental neural networks, such as the papers \cite{NN1,NN2,NN3,NN4} which use incremental neural networks to train on the data points one by one and update the weights every time after scanning a data point. The shortcomings of this incremental neural network is that the results are highly dependent on the randomized initial weights since data is only passed once. As we know, in batch learning, we pass the data in many epochs to avoid the effect of randomized initial weights. This is impossible in incremental neural networks since the data only gets passed once and previous data is discarded.

		Other researchers proposed methods based on ensemble classifiers. For example, Wang et al.\cite{ensemble1} proposed a general framework for mining data streams with concept drift using weighted ensemble classifiers. The authors train an ensemble of classification models from sequential chunks of the data stream. However, building a classifier independently on each chunk can be inefficient since the data from all chunks are usually dependent.

		Overall, most stream mining algorithms involve some stream summarization. There are different ways to summarize data streams. Some methods will only keep a portion of the data in streams by using sampling. Such approaches including Reservoir Sampling \cite{vitter}, Concise Sampling \cite{Gibbons}, Chain Sampling \cite{Babcock}, Order Statistics with sampling approximate quantiles \cite{Manku}, and Sampling from a Moving Window, such as sliding windows. Another common method for processing the stream data is filtering, such as bloom Filters. The third most common method is to construct a probabilistic summary of the data with hash functions and randomized algorithms to get approximations.

		The models proposed in this paper overcome the disadvantages of incremental neural networks and ensemble classifiers.  The SMNN algorithms learns from stream data in mini-batches with sliding windows. In each mini-batch, we run the algorithm in many epochs to reduce the effect of randomized initial weights. Moreover, we pass the weights from one mini-batch to the other mini-batch in order to obtain better weights. The method of SENN trains multiple neural networks on sliding windows and uses plurality voting on the collection of neural networks to classify new data. Moreover, whenever there is no concept drift, we pass the weights learned from a sliding window to the subsequent sliding window and use them as the initial weights of the next neural network. By doing so, the training  is much faster compared to using random weights every time. We will also show that the classifiers are diverse enough to give a good prediction.

		\section{Sliding Neural Network Algorithms} 
		\label{sec:models}
		TODO: This section isn't very clear. Should rewrite.

		As discussed in Section \ref{sec:related-work}, each stream mining algorithm usually involves some summarization method. In both of our models,  we used the \emph{sliding windows} to summarize the data streams.

		Sampling from a moving window is to maintain a \emph{sliding window} of the most recently arrived data. In other words, the window has a fixed size $L$ and it works as a first in first out queue. we chose sliding windows to summarize and sample our data, as it is simple and quite appropriate for the task of classification.

@@

		\subsection{Sliding Mini-Batch Neural Networks}
		The first model we introduce for mining a data stream is called Sliding Mini-batch Neural Networks. Essentially, this algorithm involves ``passing'' the weights of neural networks. In other words, we passed the final weights of a trained neural network on a sliding window and used them as the initial weights for training a neural network on the next sliding window.

		In our model, we combined the neural networks with sliding windows to train the data. Suppose the training dataset has size $n$. Let us fix $L$ as the size of sliding windows and $d$ as the length of the steps we slide each time, then we will have $s = \lfloor(n-L+d) / d\rfloor$ sliding windows in total. Denote the sliding windows by $Y_1, \ldots, Y_s$.  The model is trained with following steps: 
		\begin{enumerate}
			\item Randomly initialize a weight matrix $\textbf{W}_0$.
			\item Train the neural network with the $L$ data points in sliding window $Y_1$ for up to $m$ epochs. Obtain $\textbf{W}_1$ which is the final weights matrix of the neural network.
			\footnote{An epoch is a single pass through the entire training set} 
			\item Train another neural network with $\textbf{W}_1$ as initial weights on the $L$ data points in sliding window $Y_2$ for $m$ epochs. Obtain $\textbf{W}_2$, the final weights of the neural network.\\
			$\ldots$
			\item Repeat until we reach the last sliding window $Y_s$ and obtain the final weights $\textbf{W}_{s}$.
		\end{enumerate}
		A graph of the Sliding Mini-batch Neural Networks is as follows:
		\begin{figure}[H]
			\centering
			\includegraphics[width=1\linewidth]{figures/CNN}
			\caption{Sliding Mini-Batch Neural Networks}
			\label{fig:minibatchANN}
		\end{figure}
		After passing all the data in sliding windows for training, the neural network model is ready for classification. The weights matrix $\textbf{W}_{\lfloor(n-L+d) / d\rfloor}$ will be the weights of connections for the final neural network used to classify testing data.

		As we discussed in section \ref{sec:related-work}, SMNN optimizes the incremental neural networks algorithm since the initial weights are adjusted constantly by the $L$ data instances in one sliding window. By feeding the data instances in one sliding window in many epochs, the order of feeding the data becomes less important. Therefore, the performance of SMNN is less random than incremental neural networks algorithm. Therefore, SMNN has smaller bias and variance in classification.

		\subsection{Sliding Ensemble Neural Networks}
		\label{sec:senn}
		The second model we introduce is based on ensembles of neural networks with sliding windows. As shown in paper \cite{ensemble1}, ensemble of classifiers usually works better than a single classifier in mining data streams. Therefore, the main idea is to build one neural network on each sliding window and take a majority vote. However, most of the time, the data instances in data streams are dependent. For example, the temperatures measured by a sensor are usually dependent on or correlated to the last temperature measure.

		Because data instances in a data stream are usually dependent, it is natural to assume that the data instances in a sliding window are correlated to those in the previous sliding window.  Due to the dependency of data instances between different sliding windows, we concluded that it would be inefficient to build a neural network on each sliding window independently.

		We further improved the model by choosing to pass the weights from one neural network to the subsequent neural network only in the case where there is no concept drift in the dataset. Moreover, by doing so, the training time gets reduced, since passing weights provides ``good weights'' as  initialization weights for each new neural network model.

		This new classification model is called \emph{Sliding Ensemble Neural Networks} or \emph{SENN}. This model works as the following figure shows.

@@

		The building of this model is counter-intuitive because for an ensemble of classifiers to be more accurate than any of its individual members, it is necessary and sufficient for the individual classifiers to be accurate and diverse \cite{Hansen}. In this context, an \emph{accurate} classifier is one that has an error rate better than 50\%. A set of classifiers is said to be \emph{diverse} if the classification errors of the classifiers are independent from each other \cite{Dietterich}. Informally, this means that they don't make classification errors on the same new data points (more than would be allowed by chance alone). We therefore believe that for SENN, there is a tradeoff between accuracy \emph{of individual models} and diversity of the ensemble. In particular, in the case where the final weights of a neural network are passed and used as the initial weights of the subsequent neural network, that is considered equivalent to having trained each neural network on more data points than are available in a sliding window. In turns, this increases the accuracy of individual classifiers in the ensemble (in addition to reducing training time).

		Of course, the end goal is to obtain the best accuracy from ensemble voting and not from individual neural networks. Thus, it is not immediately obvious what the best approach is as there is a delicate balance to establish, which we study in more details in Section~\ref{sec:passing-weights} and Section~\ref{sec:measurement-diversity}.

		However, the diversity issue can be further improved by careful models selection. Since the size of ensembles is fixed due to memory restrictions, an algorithm for models picking is necessary and a good algorithm can improve diversity.

		\subsection{Model picking methods}
		\label{sec:model-picking}

		Models picking will also help to reduce the memory needed to run the algorithm. One of the main requirements for a stream mining algorithm was \emph{constant memory}. Models (neural networks) are added to an ensemble until its size reaches $C$. At this point, the ensemble may not be optimal yet, so we do not want to discard all further models and we need to determine a way to \emph{continue learning}. In order to do so, some models will have to be dropped from the collection so that new models can be included. The principle of models picking is that the models in the ensemble are as diverse as possible in order to give better prediction.

		Moreover, in the case where the streaming datasets we are working with admits concept drifting, it is absolutely critical	to have a good model picking method that will be able to keep up with the concept drift.

		The current model picking methods are:

		\begin{enumerate}
			\item \emph{First In, First Out.} The first method is to use a first in first out (FIFO) queue. The oldest model in the collection is dropped to make room for the new model. This is the simplest method and it allows the learning algorithm to adapt very quickly to changing concepts. However, it is susceptible to develop bias quickly if presented with a lot of unrepresentative training data.

			\item \emph{Random Replacement}. The second method involves random replacement with a non-uniform distribution. There are many non-uniform distributions that would do, the point is to make it such that models at the end of the collection rarely get replaced (long-term memory of the classifier), whereas models at the beginning of the collection frequently get replaced (ability to adapt quickly to change).

			For example, here is a valid way to do this. First, generate a random integer $r$ chosen according to the geometric distribution $\exp(\lambda = 0.1 K)$. Then, let $n = min(floor(r), K)$. We would then discard the model at position $n$ (counting from 0) in our ordered collection of models and replace it with a new model.

			\item \emph{Gramian Matrix Method.} The third method is to use the error correlation rate to pick up models in order to achieve  diversity in the ensemble. We want to replace those models who give the errors on the same data points. To decide how two models are diverse from each other, we compute the error correlation of all models on a testing dataset. For example, we use all data in one sliding window as the testing data. For each model, we print all errors made by this model. Therefore, we obtain a set of error vectors $X_1, \ldots, X_m \in \{0,1\}^n$, where $m$ is the number of models in the ensemble and $n$ is the number of data instances in the testing set. The attributes of the vectors are either 1 or 0, with 1 indicating a classification error and 0 indicating success. In order to measure the degree of diversity among the models, we then compute the pairwise dot product of all error vectors. In other words, we construct the $m \times m$ matrix $A = (a_{i,j})$ where $a_{i,j} = X_i^T X_j$. Small values of $A$ are thus considered ideal, as they indicate strong pairwise independence between models, whereas large values of $A$ are undesirable. Note that $A$ is the \emph{Gramian matrix} of $X_1, \ldots, X_m$. Models who tends to make the mistakes on the same data are undesirable in an ensemble \cite{Dietterich} and they can readily be identified from the matrix $A$.
		\end{enumerate}

@@

		\section{Experimental Design}
		\label{sec:experimental-design}
		\subsection{Dataset}
		The dataset we chose for our experiment is the Dodgers Loop Sensor Data Set from the UCI repository \cite{dataset}. It comes with two file: Dodgers.data and Dodgers.events. Dodgers.data contains 50,400 instances with 3 attributes (date, time, count of cars). Dodgers.events contains 81 instances with 6 attributes representing  Date, Begin event time, End event time, Game attendance, Away team, and W/L score.

@@

		The observations of this dataset were taken over 25 weeks with count aggregates every 5 minutes. Thus, 288 data instances were generated per day with each data instance representing the cars count in a 5 minutes slot. This dataset is typically used with the goal of predicting the presence of a baseball game at Dodgers stadium based upon traffic patterns.

@@

		Figure \ref{fig:typical-week} is the plot of a typical week of the Dodgers Loop Dataset. The blue lines indicate the beginning of games and the red lines indicates the end of games.

@@

		Instead of predicting if individual data points fall within the window of a Dodgers game (between a blue line and a red line on the same day), we chose to work on the simpler task of predicting whether or not there was a Dodgers game held on any given day.
		Therefore, we aggregate the data into days containing 288 integers each (as there are 288 five-minute intervals in a day). We then labeled days on which a Dodgers game was held with ``1'' and a day without a game as ``0''.
